---
title: "toy-data"
author: "DBlana"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load packages

No packages needed for the code below!

### Code to generate random CHI numbers

```{r CHI}

n <- 20 # this is the number of generated observations

# "sample" generates random numbers, in this case 10 numbers between 0 and 9
chi <- replicate(n, paste(sample(0:9, 10, replace=TRUE), collapse = ''))

```

### Random sex and date of birth

```{r sex-DOB}
sex <- replicate(n, sample(c('female', 'male'), 1))

dob <- as.Date(replicate(n, paste(
        sample(1950:2010, 1), '-',
        sample(c(1,3:12), 1), '-',
        sample(1:30, 1),
        sep = ''
    )))

```

### Combine into one dataframe

```{r demographics}

demographics <- data.frame(chi,
                           sex,
                           dob)

```

### Other data: Aberdeen postcodes, SIMD deciles

```{r postcode-SIMD}

postcode <- replicate(n,
                  paste(
                      'AB',
                      sample(0:5, 1),
                      sample(0:9, 1),
                      ' ',
                      sample(0:9, 1),
                      sample(LETTERS, 1),
                      sample(LETTERS, 1),
                      sep = '')
        )

decile <- replicate(n, sample(1:10,1))

```


### Create blood test dataframe with the same CHI numbers as in demographics

I've included both a numerical value and a positive/negative result

```{r blood-test}

n_tests <- 50
blood_test_results <- data.frame(
    chi = replicate(n_tests, sample(demographics$chi, 1)),
    test_date = as.Date(replicate(n_tests, paste(
        sample(1999:2020, 1), '-',
        sample(c(1,3:12), 1), '-',
        sample(1:30, 1),
        sep = ''
    ))),
    #runif generates random values from the uniform distribution with min=1 and max=10
    value = replicate(n_tests, round(runif(1, 1,10), 1)),
    # here we sample from positive/negative, with 0.2 probability of positive, 0.8 probability of negative
    result = replicate(n_tests, sample(c('positive', 'negative'), size = 1, prob = c(0.2, 0.8))))

```

### Then you can save the dataframes you want into csv files and use to test and share your code!
