---
title: "practice-with-plots"
author: "D Blana"
date: "17/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r load-packages}

library(tidyverse)
library(viridis)   # nice colour scheme
library(plotly)    # interactive visualisations

```

## Read in the data

```{r read-data}

admission_data <- read_csv("alcohol_related_admissions.csv") 
glimpse(admission_data)

```

## Median admissions

Here we calculate the median admissions for each health board and each year, using "group_by". 

```{r median-admissions}

median_admissions <- admission_data %>%
  group_by(year, health_board) %>%
  summarise(median_admissions = median(alcohol_admissions))

head(median_admissions)

```

## Simple bar chart

You can read a lot more about bar plots here: https://www.r-graph-gallery.com/barplot.html

```{r bar-chart}

median_admissions %>%
  filter(year == 2019) %>%
  rename(alcohol_admissions = median_admissions) %>%
  ggplot(aes(x=health_board, y=alcohol_admissions)) + 
  geom_bar(stat = "identity")

```

Instead of wrapping the names, let's make the graph horizontal instead using "coord_flip". Also add labels and a title.

```{r bar-horizontal}

median_admissions %>%
  filter(year == 2019) %>%
  rename(alcohol_admissions = median_admissions) %>%
  ggplot(aes(x=health_board, y=alcohol_admissions)) + 
  geom_bar(stat = "identity") +
  coord_flip() +                         
  xlab(NULL) + 
  ylab("Age-sex standardised rate per 100,000") + 
  ggtitle("Alcohol-related hospital admissions in 2019") 

```

Let's also reorder them by size.

```{r bar-reorder}

median_admissions %>%
  filter(year == 2019) %>%
  rename(alcohol_admissions = median_admissions) %>%
  ggplot(aes(x=reorder(health_board,alcohol_admissions), y=alcohol_admissions)) + 
  geom_bar(stat = "identity") +
  coord_flip() + 
  xlab(NULL) + 
  ylab("Age-sex standardised rate per 100,000") + 
  ggtitle("Alcohol-related hospital admissions in 2019") 

```

Exercise 1: Could you make Grampian stand out by giving it a different colour?

```{r bar-Grampian}

median_admissions %>%
  filter(year == 2019) %>%
  rename(alcohol_admissions = median_admissions) %>%
  mutate(isGrampian=ifelse(health_board=="Grampian","yes","no")) %>%
  ggplot(aes(x=reorder(health_board,alcohol_admissions), 
             y=alcohol_admissions,
             fill=isGrampian)) + 
  geom_bar(stat = "identity") +
  coord_flip() + 
  xlab(NULL) + 
  ylab("Age-sex standardised rate per 100,000") + 
  ggtitle("Alcohol-related hospital admissions in 2019") + 
  theme_bw() +
  theme(legend.position = "none")                                
  
```

Let's make a lollipop graph instead.(https://www.r-graph-gallery.com/lollipop-plot.html) 

```{r lollipop}

median_admissions %>%
  filter(year == 2019) %>%
  rename(alcohol_admissions = median_admissions) %>%
  ggplot(aes(x=reorder(health_board,alcohol_admissions), 
             y=alcohol_admissions)) + 
  geom_segment(aes(xend=health_board, y=0, yend=alcohol_admissions)) +
  geom_point(size=4, alpha=0.6) +
  coord_flip() + 
  xlab(NULL) + 
  ylab("Age-sex standardised rate per 100,000") + 
  ggtitle("Alcohol-related hospital admissions in 2019") + 
  theme_bw() 

```

Bar chart with two groups, for years 2010 and 2019

```{r grouped2-barchart}

median_admissions %>%
  filter(year %in% c(2010, 2019)) %>%   # only select these two years
  mutate(year = as.factor(year)) %>%          # convert years to categories ("factors")
  rename(alcohol_admissions = median_admissions) %>%
  ggplot(aes(x=health_board,
         y=alcohol_admissions,
         fill = year)) +
  geom_bar(position="dodge", stat = "identity") +  # without "dodge", we get a stacked bar chart
  coord_flip() + 
  xlab(NULL) + 
  ylab("Age-sex standardised rate per 100,000") + 
  ggtitle("Alcohol-related hospital admissions") + 
  theme_bw()  

```

Exercise 2: Now try making a bar chart with three groups three years 2010, 2015 and 2019.
Could you also improve the colours by using a different colour scheme?

(Hint: viridis includes a few different scales, which you can set using the "option" parameter in scale_fill_viridis, run ?scale_fill_viridis to see all options.You may also want to add a "begin" or "end" parameter, to make sure the lightest colour is not so light it's not visible against the white background.)

```{r grouped3-barchart}

median_admissions %>%
  filter(year %in% c(2010, 2015, 2019)) %>%   # only select these three years
  mutate(year = as.factor(year)) %>%          # convert years to categories ("factors")
  rename(alcohol_admissions = median_admissions) %>%
  ggplot(aes(x=health_board,
         y=alcohol_admissions,
         fill = year)) +
  geom_bar(position="dodge", stat = "identity") +  # without "dodge", we get a stacked barchart
  coord_flip() + 
  xlab(NULL) + 
  ylab("Age-sex standardised rate per 100,000") + 
  ggtitle("Alcohol-related hospital admissions") + 
  scale_fill_viridis(discrete=TRUE, option="magma",end = 0.9)+ 
  theme_bw()  

```

## Is there a relationship between alcohol-related hospital admissions and deprivation?

Let's load the second dataset, "simd20_2019.csv". This contains population estimates for each SIMD decile per health board in Scotland for 2019. It came from here: https://www.nrscotland.gov.uk/statistics-and-data/statistics/statistics-by-theme/population/population-estimates/2011-based-special-area-population-estimates/population-estimates-by-simd-2016

```{r read-deprivation-data}

deprivation_data <- read_csv("simd20_2019.csv") 
glimpse(deprivation_data)

```

Let's calculate the percent of the population in the most deprived areas (SIMD deciles 1-3) in each health board.

In the code below, we use first mutate & sum and then summarise & sum to calculate the sum of a column. The difference is that mutate adds a column to the existing dataframe, while summarise creates a new aggregate dataframe. Read more about this here: https://argoshare.is.ed.ac.uk/healthyr_book/summarise-vs-mutate.html 

```{r percent-pop-simd}

# Calculate total population per health board, and population per SIMD decile, then divide to get the percent of the population at each SIMD decile

deprivation_data <- deprivation_data %>%
  group_by(`NHS health board`) %>%
  mutate(Total_population = sum(Population),
         Percent_population = Population/Total_population) %>%   # here we are adding new columns to the original dataframe
  ungroup()

# We can now keep SIMD = 1-3

deprived_ratio <- deprivation_data %>%
  filter(`SIMD decile` <= 3) %>%
  group_by(`NHS health board`) %>%
  summarise(percent_deprived = sum(Percent_population))  # here we are creating a new "summary" dataframe

```

Exercise 3: Link this dataset to median_admissions for 2019. (Remember to only keep the median alcohol admission data for the year 2019.)

```{r join-admissions-deprivation}

median_admissions_2019 <- median_admissions %>%
  filter(year == 2019)

admissions_deprivation_2019 <- left_join(median_admissions_2019,deprived_ratio,by = c("health_board" = "NHS health board"))

glimpse(admissions_deprivation_2019)

```

Exercise 4: Now create a scatterplot looking at the relationship between percent deprived and alcohol-related hospital admissions.

```{r scatterplot}

admissions_deprivation_2019 %>% 
  ggplot(aes(x = percent_deprived, 
             y = median_admissions)) +       
  geom_point() +
  xlab("Percent of the population living in SIMD deciles 1-3") +            
  ylab("Age-sex standardised rate per 100,000") +              
  ggtitle("Alcohol-related hospital admissions in 2019 across health boards") + 
  theme_bw() 

```

Possible improvements:
1. Add text (https://ggplot2.tidyverse.org/reference/geom_text.html)

```{r add-text}

admissions_deprivation_2019 %>% 
  ggplot(aes(x = percent_deprived, 
             y = median_admissions)) +       
  geom_point() +
  geom_text(aes(
    label = health_board,
    check_overlap = TRUE)) +
  xlab("Percent of the population living in SIMD deciles 1-3") +            
  ylab("Age-sex standardised rate per 100,000") +              
  ggtitle("Alcohol-related hospital admissions in 2019 across health boards") + 
  theme_bw() 

```

2. Make your visualisation interactive

```{r interactive}

p <- admissions_deprivation_2019 %>% 
  ggplot(aes(x = percent_deprived, 
             y = median_admissions,
             label = health_board)) +       
  geom_point() +
  xlab("Percent of the population living in SIMD deciles 1-3") +            
  ylab("Age-sex standardised rate per 100,000") +              
  ggtitle("Alcohol-related hospital admissions in 2019 across health boards") + 
  theme_bw()

ggplotly(p,tooltip = "label")

```


3. Turn it into a bubble plot, using the total population of each health board for the circle size (https://www.r-graph-gallery.com/320-the-basis-of-bubble-plot.html). You will need to add the total population of each health board to the dataframe.

```{r bubble-plot}

# We need to add the total population to our dataframe

pop_data <- deprivation_data %>%
  select(`NHS health board`,Total_population) %>%
  distinct()

adm_deprivation_2019 <- left_join(admissions_deprivation_2019,pop_data,by = c("health_board" = "NHS health board"))

p <- adm_deprivation_2019 %>% 
  ggplot(aes(x = percent_deprived, 
             y = median_admissions,
             label = health_board,
             size = Total_population)) +       
  geom_point(alpha=0.7) +
  xlab("Percent of the population living in SIMD deciles 1-3") +            
  ylab("Age-sex standardised rate per 100,000") +              
  ggtitle("Alcohol-related hospital admissions in 2019 across health boards") + 
  theme_bw() +
  theme(legend.position = "none") 

ggplotly(p,tooltip = "label")

```







