---
title: "final R practice"
author: "JW"
date: "2024-11-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load packages
```{r}
library(tidyverse)
```

## read in data
```{r}
data_hospital_admissions <- read.csv('hospital_admissions.csv')
data_population_demographics <- read.csv('population_demographics.csv')

summary(data_hospital_admissions)

summary(data_population_demographics)

tail(data_hospital_admissions, 10)

tail(data_population_demographics, 10)
```

### practice 1.1) any problems with the values of age and deprivation?

```{r}
data_population_demographics$age <- as.numeric(data_population_demographics$age)

arranged_population_demographics <- arrange(data_population_demographics, by= age)

head(arranged_population_demographics)

age_0 <- sum(arranged_population_demographics$age==0)

print(age_0)

table(arranged_population_demographics$age==0)

n_na<- sum(is.na(arranged_population_demographics$deprivation))

print(n_na)
```

There are 'r age_0' people in population_demographics whose age is 0. 

There are 'r n_na' people in population_demographics whose deprivation is NA.

##remove age=0 and remove NA in deprivation from population_demographics

```{r}
cleaned_population_demographics <- arranged_population_demographics  %>% 
  filter(age!=0,!is.na(deprivation)) 

duplicate_id<- cleaned_population_demographics %>% 
  group_by(id) %>% 
  filter(n()>1)

print(duplicate_id)

head(cleaned_population_demographics)

```

### practice 1.2) minimum, maximum and median age for each deprivation level?

```{r}
age_deprivation <- cleaned_population_demographics %>% 
  group_by(deprivation) %>% 
  summarise(min=min(age),max=max(age),median=median(age))

head(age_deprivation)
```


### practice 2.1) create a data frame that contains both demographic and hospital admission data
```{r}
join_demographic_admission<- full_join(data_hospital_admissions,data_population_demographics) %>% 
  filter(age!=0,!is.na(deprivation)) 

summary(join_demographic_admission)
```

### practice 2.2) summarise total time in hospital in 2020 by age and deprivation. Include a graph you think would be useful to explain your answer to a general audience

```{r}
head(join_demographic_admission)
```


