---
title: "Practical Data Cleaning with R"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

this practice is from PU5058, PU5063 (2024 -25): Introduction to Health Data Science Practical: Data Cleaning with R week 6

there are 3 questions

1. Plot 'EQ-5D Index' scores pre and post operation for each gender

2. Calculate how many patients in this dataset have been told by a doctor that they have problems caused by a stroke

3. Create a clean and tidy table with pre and post operation activity levels

### Load packages

```{r}
library(tidyverse)
```

### First question, Plot 'EQ-5D Index' scores pre and post operation for each gender

## read in data and see the column names

```{r read in data}
raw_hip_data <- read_csv('Hip Replacement CCG 1819.csv')

head (raw_hip_data)

names(raw_hip_data)
```

##remove NA

```{r remove NA}
raw_hip_data_noNA <- raw_hip_data %>% 
  drop_na() %>% 
  filter(Gender!='*') %>% 
  rename(EQ5D_Index_PreOp='Pre-Op Q EQ5D Index',EQ5D_Index_PostOp='Post-Op Q EQ5D Index')

head(raw_hip_data_noNA)
```

## select Gender, 'EQ-5D Index' scores pre and post operation. replace 1=male(M), 2= female(F) in Gender column. add patient ID before plotting.

```{r select 3 colomns}
Gen_EQ5D_Index <- raw_hip_data_noNA %>% 
  select('Gender','EQ5D_Index_PreOp','EQ5D_Index_PostOp') %>% 
  mutate(Gender = recode(Gender, `1` = "M", `2` = "F")) %>% 
  mutate(patient_ID = row_number())

head(Gen_EQ5D_Index)
summary(Gen_EQ5D_Index)
```

## Plot 'EQ-5D Index' scores pre and post operation for each gender

```{r}
tidy_Gen_EQ5D_Index <- Gen_EQ5D_Index %>%
  pivot_longer(
    c('EQ5D_Index_PreOp','EQ5D_Index_PostOp'),
    names_to= 'Time',
    names_prefix='EQ5D_Index_',
    values_to= 'EQ5D_Index'
    )

head(tidy_Gen_EQ5D_Index)
summary(tidy_Gen_EQ5D_Index)
```

## change order of Time and visualization of the frame!
```{r}

tidy_Gen_EQ5D_Index$Time <- factor(tidy_Gen_EQ5D_Index$Time, levels=c('PreOp','PostOp'))

tidy_Gen_EQ5D_Index %>%
  
  ggplot() +
  geom_boxplot(aes(x = Time, y = EQ5D_Index, colour = Gender)) 

```


## Try the transformation to normal (use log)
```{r}
EQ5D_Index_log <- tidy_Gen_EQ5D_Index %>% 
   mutate(EQ5D_Index_log =log(EQ5D_Index))

tidy_Gen_EQ5D_Index$Time <- factor(tidy_Gen_EQ5D_Index$Time, levels=c('PreOp','PostOp'))

EQ5D_Index_log %>%
  
  ggplot() +
  geom_boxplot(aes(x = Time, y = EQ5D_Index_log, colour = Gender)) 

```


### Second, calculate how many patients in this dataset have been told by a doctor that they have problems caused by a stroke

```{r filter patients with stroke problems}
patient_stroke <- raw_hip_data_noNA %>% 
  rename(heart_disease='Heart Disease') %>% 
  filter(Stroke==1) 

count(patient_stroke)
n1 <- nrow(patient_stroke)
```

There are `r n1` patients in this dataset have been told by a doctor that they have problems caused by a stroke.

### Third, create a clean and tidy table with pre and post operation activity levels

```{r select pre and post op activity levels}
pre_and_post_act <- raw_hip_data_noNA %>% 
  select('Gender','Pre-Op Q Activity', 'Post-Op Q Activity') %>% 
  mutate(patient_ID=row_number())
head(pre_and_post_act)
summary(pre_and_post_act)
```

```{r tidy table with pre and post op activity levels}
tidy_pre_post_act <- pre_and_post_act %>% 
  rename(PreOp='Pre-Op Q Activity',PostOp='Post-Op Q Activity') %>% 
  pivot_longer(c(PreOp,PostOp),
               names_to='Time',
               values_to='Activity')
head(tidy_pre_post_act)
summary(tidy_pre_post_act)
```