---
title: "alcohol-admins_deprive-practice2"
author: "JW"
date: "2024-11-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Exercise 2: For which health boards do we have data on alcohol-related hospital admissions for each deprivation quintile? 


## Load packages

```{r}
library(tidyverse)
```

## read in data

```{r}
depri_raw_data <- read_csv("alcohol-admissions_deprivation-data.csv") 

inter_zone_codes <- read_csv("iz2011_codes_and_labels_21042020.csv") 

unique(depri_raw_data$geography_code)

```

## filter data and select geography_code
```{r}

depri_health_board <- filter(inter_zone_codes, HB == 'S08000020' |   HB == 'S08000029') %>% 
  select(HB, HBName)

table(depri_health_board)
```


Exercise 3: Create boxplots showing alcohol admissions for each deprivation quintile. Create a different plot for each health board in the data set. 

##prepare data

```{r}
depri_quintile <- depri_raw_data %>% 
  inner_join(inter_zone_codes, by=c('geography_code'='HB')) %>%
  select(HBName,quintile,indicator_measure) %>% 
rename(health_board='HBName',admissions='indicator_measure') %>% 
aggregate(health_board ~ quintile, data = depri_raw_data, FUN = sum)

table(depri_quintile)
```


```{r}
getwd()

write_csv(depri_quintile,"alcohol_admissions_deprive.csv") ##where is it?
```


##box-plot alcohol admissions for each deprivation quintile
```{r}
depri_quintile %>%                          
  ggplot(aes(x = quintile,             
             y = admissions)) +    
  geom_boxplot() +               
 ggtitle("Alcohol Admissions for Each Deprivation Quintile") +
facet_wrap(~health_board)+
   theme(legend.position = "top")  #how to show legend 1=most deprived...
```


