---
title: "alcohol-admins_deprive-practice"
author: "JW"
date: "2024-11-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Exercise 2: For which health boards do we have data on alcohol-related hospital admissions for each deprivation quintile? 


## Load packages

```{r}
library(tidyverse)
```

## read in data

```{r}
depri_raw_data <- read_csv("alcohol-admissions_deprivation-data.csv") 

inter_zone_codes <- read_csv("iz2011_codes_and_labels_21042020.csv") 

unique(depri_raw_data$geography_code)

```

## filter data and show health boards in deprive table
```{r}
inter1 <- inter_zone_codes %>% 
  select(HB, HBName) %>% 
  group_by(HB, HBName) %>% 
  summarize

depri_health_board <- depri_raw_data %>% 
                      left_join(inter1,                                                                              by = c('geography_code' = 'HB')) %>% 
  select(geography_code, HBName,quintile,indicator_measure)

unique(depri_health_board$HBName)

```
We have two health boards--`unique(depri_health_board$HBName)`-- data on alcohol-related hospital admissions for each deprivation quintile.

Exercise 3: Create boxplots showing alcohol admissions for each deprivation quintile. Create a different plot for each health board in the data set. 

##prepare data

```{r}
depri_quintile <- depri_health_board %>% 
                  rename(health_board='HBName',admissions='indicator_measure') 

write_csv(depri_quintile,'depri_health_board.csv')

glimpse(depri_quintile)

```

##box-plot alcohol admissions for each deprivation quintile
```{r}
depri_quintile %>%                          
  ggplot(aes(x = quintile,             
             y = admissions)) +    
  geom_boxplot() +               
  ggtitle("Alcohol Admissions for Each Deprivation Quintile") +
  facet_wrap(~health_board) +
  theme(legend.position = "top")  #how to show legend 1=most deprived...
```


